Microlensing as a possible probe of event-horizon structure in quasars

### 2017-06-26 paper1/ms.tex
editing

%In Section~\ref{sec:source-models} we introduce the three source
%profiles used in our subsequent models and simulations: a
%constant-brightness disc, a circular Gaussian, and the crescent source
%introduced by \citep{2013MNRAS.434..765K}.  The latter is simply a
%constant-brightness disc with a smaller, non-concentric disc cut out
%of it.  We also derive the half-light radius for a crescent.  The
%half-light radius can characterise the source size for all three types
%of source.

In Section~\ref{sec:numerics} we carry out source fitting to
lightcurves, with both noise and systematic errors are present.  We
generate a mock lightcurves by taking a crescent source across the
path AB in figure~\ref{fig:magnification_map}, and then adding noise.
The path simulates crossing a clean but not ideal fold.  We then fit
this lightcurve to templates from both crescent and circular-Gaussian
sources across the path CB.  That is, the templates come from a
similar but not identical caustic, thus deliberately generating a
systematic error.  Despite this systematic error, a crescent source
fits the data, while a Gaussian source is rejected from the $\chi^2$
value.

@@

In Section~\ref{sec:numerics} we carry out source fitting to
lightcurves, with both noise and systematic errors are present.  We
generate three lightcurves by taking a uniform disc, a circular
Gaussian, and a crescent across the path AB in
Figure~\ref{fig:magnification_map}, and then adding noise.  The path
simulates crossing a clean but not ideal fold.  In addition, we
generate templates lightcurves by running the three source types, with
various parameter values, across the path CB.  That is, the templates
come from a similar but not identical caustic, thus deliberately
generating a systematic error.  We then fit the noisy lightcurves to
the templates using Markov chain Monte-Carlo.  We find that the
correct source type can be inferred from the $\chi^2$ values.  The
parameter values can also be inferred. The fitting errors are larger
than the formal uncertainties, which is expected in the presence of
systematic errors, but still appear acceptable.

@@

Once the map is created, the lightcurve can be obtained by specifying
the transit path of the source across the map.  The source is moved
along the transit line, one pixel at a time, and at each step the
brightness distribution of the source is convolved
(equation~\ref{eqn:ft2d}) with the magnification map to give the
brightness.  In real life, not only are both lens and source moving
but the lens configuration, and with it the magnification
pattern, is also changing with time.  While the first subtlety is
taken care of by a coordinate transformation in this analysis, for the
second one the lens configuration is assumed to be constant in time.

For the purpose of this analysis, it was desirable to mimic the
analytical behaviour of a simple fold as much as possible. For
comparing the numerical result with the analytical one, therefore, the
path of the source was chosen, so that it intersects the border of the
caustic perpendicularly, and on a point where the border is a fold
caustic.  We choose the line AC in Figure~\ref{fig:magnification_map}
and consider its two caustic-crossing segments separately:
\begin{itemize}
\item ``path~I'' runs from C to B;
\item ``path II'' runs from A to B.
\end{itemize}
The magnification along paths I and II are very similar, but not
identical.

The mock lightcurve used a crescent source (as explained in
\S\ref{subsec:crescent}) with parameters
\begin{equation}
   \left(R_p, \frac{R_n}{R_p}, \frac{a}{R_p}, \frac{b}{R_p}\right) =
   (1.31, 0.4, 0, 0.3)
\label{eqn:cp}
\end{equation}
corresponding to $r_{1/2}=1$.  Unit length has 30 pixels on the
magnification map.  The source was carried for 250 pixels along part
of path~I, giving 251 brightness values.  Gaussian noise at the level
of 4\% of the current brightness was then added to give the mock
lightcurve.  If we assume the brightness measurements are nightly, and
the transverse speed is 600~km/s, $r_{1/2}$ comes to 24~light-minutes,
which is plausible for a quasar.

@@

@@

%In the case of a high-quality lightcurve with 
%insignificant noise and measurement errors, 
%the parameters can be obtained 
%by simply identifying the characteristic 
%instances of time without making use of 
%the actual values of the magnification
% map. If the effect of the errors and the 
%noise distorts the magnification time function enough so that the 
%characteristic epochs are not identifiable with the characteristic 
%periods still visible, the boundaries of the periods 
%can be roughly estimated. Furthermore, if direct estimates of 
%the parameters cannot be obtained we propose the use of a strong
%statistical tool such as Markov-Chain Monte Carlo.
% In Section 6  we have studied the possibility of identifying a crescent 
%source and the possibility of recovering the 
%respective parameters. As magnification map, we have used a complex numerical 
%one generated with the microlensing code by \cite{1999A&A...346L...5W}. 
%In addition, we have added to the signal a Gaussian 
%noise with an SNR of 1.6.  In the experiment,
% we have considered all nine combinations of
% original source profiles and assumed fitting source models.
%Effectively we have fitted using MCMC all three sources with all three assumed fitting models. The results of the experiment allowed us to build a 
%procedure for distinguishing the shape of the source assuming that one of the three models we have considered is a good approximation.  
%The procedure would be useful for observers that endeavour to gain more information about an unresolved source for which they can study 
%the microlensing lightcurve.  As a first step in the procedure, one should first attempt to use MCMC with a Gaussian model assumption. 
%If either the value of the $\chi^2$ or the number of rejected datapoints is large then the next step is to change the assumed source model to 
%a uniform disc and redo the fitting. Finally, if the uniform disc assumption is rejected as well then the fitting should be done with a 
%crescent source model assumption. If at each of the three steps the data rejects the model then the source cannot be approximated by any of the three models. 
%Otherwise, if at one of the steps the data does not reject the model then the respective model is a good approximation.

>>>>

In Section~\ref{sec:source-models} we introduce the three source
profiles used in our subsequent models and simulations: a
constant-brightness disc, a circular Gaussian, and the crescent source
introduced by \citep{2013MNRAS.434..765K}.  The latter is simply a
constant-brightness disc with a smaller, non-concentric disc cut out
of it.  We also derive the half-light radius for a crescent.  The
half-light radius can characterise the source size for all three types
of source.

@@

In Section~\ref{sec:numerics} we carry out source fitting to
lightcurves, with both noise and systematic errors are present.  We
generate a mock lightcurves by taking a crescent source across the
path AB in figure~\ref{fig:magnification_map}, and then adding noise.
The path simulates crossing a clean but not ideal fold.  We then fit
this lightcurve to templates from both crescent and circular-Gaussian
sources across the path CB.  That is, the templates come from a
similar but not identical caustics, thus deliberately generating a
systematic error.  Despite this systematic error, a crescent source
fits the data, while a Gaussian source is rejected from the $\chi^2$
value.

@@

Once the map is created, the lightcurve can be obtained by specifying
the transit path of the source across the map.  The source is moved
along the transit line, one pixel at a time, and at each step the
brightness distribution of the source is convolved
(equation~\ref{eqn:ft2d}) with the magnification map to give the
observable total brightness.  In real life, not only are both lens and
source moving but the lens configuration, and with it the
magnification pattern, is also changing with time.  The first subtlety
is taken care of by a coordinate transformation in this analysis; the
second one is not considered, and the lens configuration is assumed to
be constant in time.

For the purpose of this analysis, it is desirable to mimic the
analytical behaviour of a simple fold as much as possible.
Accordingly, we generated a mock lightcurve by taking a crescent
source along the path CB in Figure~\ref{fig:magnification_map}.  The
crescent parameters (as defined in \S\ref{subsec:crescent}) were
\begin{equation}
   \left(R_p, \frac{R_n}{R_p}, \frac{a}{R_p}, \frac{b}{R_p}\right) =
   (1.31, 0.4, 0, 0.3)
\label{eqn:cp}
\end{equation}
corresponding to $r_{1/2}=1$.  Unit length has 30 pixels on the
magnification map.  The source was carried for 250 pixels along the
path CB, giving 251 brightness values.  Gaussian noise at 4\% of the
current brightness was then added to give the mock data.

The time unit for the light curve is taken to be $r_{1/2}/v_{\!\perp}$
where $v_{\!\perp}$ is the velocity transverse to the caustic.  If we
assume the brightness measurements are nightly, and
$v_{\!\perp}=\rm200\,km/s$, the length of the microlensing event (250
days) is comparable to typical observed events \cite[see
  e.g.,][]{2015ApJ...814L..26M}. The implied $r_{1/2}$ is
$\sim10^{14}\rm\,cm$, which is at the lower end of values inferred for
microlensed X-ray sources, whereas for optical sources, the inferred
$r_{1/2}$ is one to two order of magnitude larger \citep[see, e.g.,
  Figure~8 of][]{0004-637X-769-1-53}.

The mock data appear in Figures
\ref{fig:cc_forward}--\ref{fig:gc_forward} below.  The example values
abovr for $r_{1/2}$, $v_{\!\perp}$ and the observing frequency are not
important.  What is important is that the observations have enough
sampling and signal-to-noise that the characteristic features of a
crescent source (see e.g., Figure~\ref{fig:char_points}) are not
washed out.

@@

Widely used \citep[e.g.,][]{2010ApJ...712..658P}.

@@
